Mergeable Types

### 2017-02-27 paper/evaluation.tex

While updates are propagated asynchronously, we also measured the impact of
propagating updates synchronously by forcing synchronization every few
operations. This corresponds to a partially synchronous system, which is also
unavailable; if the cluster is unreachable, local operations will not be
accepted. While this is not a realistic deployment, the results help us
understand the overhead of synchronization. The results are shown in
Figure~\ref{grf:synchronization}. The result show that frequent blocking
synchronization is prohibitively expensive with the throughput dropping to 3
ops/s for synchronizing after every operation. \name uses Git transfer protocol
for synchronization which involves multiple roundtrips between the nodes for
synchronization. But as the synchronizations get less frequent, we can see that
the throughput asymptotically approaches the asynchronous case
(Figure~\ref{grf:scalability}).

Finally, we evaluate the performance of merge functions. For this experiment,
we consider the nodes 1 (\C{us-central1-b}) and 2 (\C{europe-west1-b}). Both
nodes start with the initial document and perform \C{n} concurrent edits
without synchronization. After this, the nodes synchronize pushing the local
edits and pulling the remote edits, and merge to produce the final document. In
this case, we indicate the distance of synchronization as \C{2n}.
Figure~\ref{grf:merge} presents the time share between network and computation
for performing 3-way merge. We see that the merge time is dominated by the
network overhead. This is because of the transfer protocol involving multiple
roundtrips and the need to transfer objects not only for application objects
but also Git internal objects for history, directory structure and branches. We
observe that we can improve the network performance by implementing a streaming
protocol that eagerly transfers objects.

More importantly, the cost of computation grows sub-linearly. This is because
of efficient merge function on ropes enabled by Irmin's content-addressed
storage backend, and more expensive operational transformation of merging
concurrent edits to leaf nodes is invoked rarely on small strings.

>>>>

While updates are propagated asynchronously, we also measured the impact of
propagating updates synchronously by forcing synchronization every few
operations. This corresponds to a partially synchronous system, which is also
unavailable; if the cluster is unreachable, local operations will not be
accepted. While this is not a realistic deployment, the results help us
understand the overhead of synchronization. The results are shown in
Figure~\ref{grf:synchronization}. The result show that frequent blocking
synchronization is prohibitively expensive with the throughput dropping to 3
ops/s for synchronizing after every operation. \name uses Git transfer protocol
for synchronization which involves multiple round trips between the nodes for
synchronization. But as the synchronizations get less frequent, we can see that
the throughput asymptotically approaches the asynchronous case
(Figure~\ref{grf:scalability}).

Finally, we evaluate the performance of merge functions. For this experiment,
we consider the nodes 1 (\C{us-central1-b}) and 2 (\C{europe-west1-b}). Both
nodes start with the initial document and perform \C{n} concurrent edits
without synchronization. After this, the nodes synchronize pushing the local
edits and pulling the remote edits, and merge to produce the final document. In
this case, we indicate the distance of synchronization as \C{2n}.
Figure~\ref{grf:merge} presents the time share between network and computation
for performing 3-way merge as we increase the operation distance between the
remote branches. We see that the merge time is dominated by the network
overhead. This is because of the transfer protocol involving multiple
round trips and the need to transfer objects not only for application objects
but also Git internal objects for history, directory structure and branches. We
observe that we can improve the network performance by implementing a streaming
protocol that eagerly transfers objects.

More importantly, the cost of computation grows sub-linearly. This is because
of efficient merge function on ropes~\ref{ropes} enabled by Irmin's
content-addressed storage backend, and more expensive operational
transformation of merging concurrent edits to leaf nodes~\ref{lists} is invoked
rarely on small. On average, we measured that at a distance of 256, for merging
the two documents with 1743 characters, 50 list merges were performed on
strings of average length 13. A purely list based approach would require the
polynomial time edit sequence algorithm to be applied on the entire document
twice. Compare to this, the rope implementation has much better algorithmic
complexity. We measured that by representing the entire document using
mergeable list type, in addition to slowing down each edit operation, the merge
at a distance of 256 was 2.3 times slower than the rope implementation.

### 2017-02-27 paper/intro.tex

The remainder of the paper is organized as follows.  We present the
\name\ programming model informally in the next section.
Sec.~\ref{sec:merge-examples} develops a series of examples that
illustrate how merge operations can be defined and used.
Sec.~\ref{sec:formalization} defines an operational semantics, and
formalizes our intution of a mergeable type in terms of morphisms over
datatype operations.  We also present soundness (consistency) and
progress guarantees enjoyed by the
semantics. Sec.~\ref{sec:system-model} considers an instantiation of
the model suitable for distributed environments with network
partitions and failures.  Experimental results

>>>>

The remainder of the paper is organized as follows.  We present the
\name\ programming model informally in the next section.
Sec.~\ref{sec:mergeable_types} develops a series of examples that
illustrate how merge operations can be defined and used.
Sec.~\ref{sec:formalization} defines an operational semantics, and
formalizes our intution of a mergeable type in terms of morphisms over
datatype operations.  We also present soundness (consistency) and
progress guarantees enjoyed by the
semantics. Sec.~\ref{sec:system-model} considers an instantiation of
the model suitable for distributed environments with network
partitions and failures.  Implementation details describing how we
incorporate mergeable types into OCaml are presented in
Sec.~\ref{sec:impl}.  Experimental results and benchmarks are given
in Sec.~\ref{sec:evaluation}.
